Structured Genealogy Processes

### 2022-01-28 ms.Rnw
sgp results updated

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=1.8]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (A) at (0,0) {$\Master_t$};
      \node (X) at (180:2) {$\X_t$};
      \node (H) at (216:2) {$\Hist_t$};
      \node (I) at (252:2) {$\Inv_t$};
      \node (G) at (288:2) {$\Gen_t$};
      \node (P) at (324:2) {$\Prune_t$};
      \node (V) at (360:2) {$\Vis_t$};
      \node (W) at (342:1.5) {$\Wch_i$};
      \node (O) at (306:1.2) {$\Och_i$};
      \draw [det,->] (A) -- (X);
      \draw [det,->] (A) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (A) -- (I);
      \draw [det,->] (A) -- (G);
      \draw [det,->] (A) -- (V);
      \draw [det,->] (A) -- (P);
      \draw [det,->] (A) -- (O);
      \draw [det,->] (P) -- (O);
      \draw [det,->] (A) -- (W);
      \draw [det,->] (V) -- (W);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [stoch,->] (H) -- (I);
      \draw [stoch,->] (H) -- (G);
      \draw [stoch,->] (H) -- (V);
      \draw [stoch,->] (H) -- (P);
      \draw [stoch,->] (H) -- (O);
      \draw [stoch,->] (H) -- (W);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

@@

\paragraph{Transitions}
Infective demes are interchangeable and we define the transition function from Deme $i$ to $j$ for any $i\neq j$ and $i,j=1,\dots,J$ as $M_{ij}:\Uspace\to\N$ to represent the number of infectives transiting to Deme $j$ from Deme $i$.
Thus the set or transition is $\Mset_{ij}\coloneqq\,M^{-1}_{ij}(\{1\})$ since we assume single transition at a single transition event. 
Similarly, we disallow coninciding transition events and other events.

@@

\paragraph{Deaths and population size}
The population, the set of infected individuals in the context of epidemiology, is structured, categorized into interchangeable demes. 
We can then define population size function and death functions in Deme $j$, respectively, as $I_{j}, D_{j}:\Xspace\to\N$ for any $j=1,\cdots,J$. 
\begin{equation}
  \jumprate_u(t,x) > 0 \Longrightarrow I_j(x+u)-I_j(x)=\sum_{i}\,B_{ij}(u)+\sum_{i}M_{ij}(u)-D_j(u)-\sum_{i}M_{ji}(u),
\end{equation}

for all $x, u\in\Xspace$. We disallow simultaneous death at a single death event at this stage and define the set of death events in Deme $j$ as $\Dset_j\coloneqq\,D^{-1}(\{1\})$. 
We also insist, birth and death events don't coincide for any demes, $i,j$, that is,  $\Bset_{j}\cap\Dset_{i}=\emptyset$. 
Eventually, we can have the function for total population size by letting $I(x)\coloneqq\sum_{j=1}^J\,I_j(x)$ and it's trivial that

@@

We can customize the state and event vector to fit a complex system. 
In SI$^2$R model, we have two different while interchangeable demes of infectious individuals, the \emph{per capita} transmissibility therefore in the first deme is $b_1(t)$ whilst in the second deme is $b_2(t)$. 
Transition rates between Deme $i$ and $j$ are also defined: Deme $i$ infection turns into Deme $j$ at rate $r_{ij}$ for $i,j=1,2$ and $i\neq j$. Furthermore, we also suppose that, once being infected, a susceptible individual becomes Deme 1 infectious with probability $\rho$ while Deme 2 infectious with probability $1-\rho$. 
To specify each event and the number of individuals involved, we define a relatively complex state vector by setting $d=10$ and $\sg{X}=\left(s,i_1,i_2,b_{11},b_{12},b_{21},b_{22},m_{12},m_{21},g\right)$, for $s$, $i_1$, and $i_2$ being the population size of the susceptible, the infectives in Deme 1, and those in Deme 2, respectively, and $b_{11}$, $b_{12}$, $b_{21}$, $b_{22}$, $m_{12}$, and $m_{21}$ being the cumulative number of events of parents in Deme 1 giving to children in Deme 1 and 2, parents in Deme 2 giving birth to children in Deme 1 and 2, individuals transiting from Deme 1 to Deme 2, and vice versa, and sampling, respectively. 
In this case, we can summarize the event rates as:

@@

The nine non-zero rates above represent those of infection into Deme 1 infected by individuals in Deme 1, infection into Deme 2 infected by individuals in Deme 1, infection in Deme 1 infected by individuals in Deme 2, infection into Deme 2 infected by individuals in Deme 1, transition from Deme 1 to Deme 2, transition from Deme 2 to Deme 1, recovery from Deme 1, recovery from Deme 2, and sampling, respectively. 
We then summarize the population functions and sets as follows:

@@

\paragraph{Migration}
One special case of the SI$^2$R model is that with one-time migration or introduction. 
In other words, an infectious individual from another population join our interested population at a certain point of time $\tau$ and then stay and spread the disease. 
In this case, if this specific individual is defined as Deme 1 while other infectious individuals as Deme 2, the transition rate $r_{12}$ is a Dirac delta function $\delta(t)$, where $\delta(t)=\infty$ for $t=\tau$ and $\delta(t)=0$ for $t\neq\tau$, and $r_{21}=0$ all the time.

@@

\begin{equation}
  \mathrm{inven}(\omega,j)\coloneqq
  \begin{cases}
    \mathrm{initialize}(I_j(\sg{X}_0(\omega)), & \text{if }K(\omega)=0; \\
    \mathrm{add}(\mathrm{inven}(\predec{\omega},j)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\Bset_j;\\
    \mathrm{drop}(\mathrm{inven}(\predec{\omega},j),N_{K(\omega)}(\omega)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\Dset_j; \\
    \mathrm{movein}(\mathrm{inven}(\predec{\omega},j),N_{K(\omega)}(\omega)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\bigcup_i\Mset_{ij}; \\
    \mathrm{drop}(\mathrm{inven}(\predec{\omega},j),N_{K(\omega)}(\omega)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\bigcup_i\Mset_{ji}; \\
    \mathrm{inven}(\predec{\omega},j), & \text{otherwise}.
  \end{cases}
\end{equation}

@@

\paragraph{Structured genealogies}
Types of nodes are defined in a structured genealogy: \emph{Leaves} are the extant members at current time and \emph{Internal Nodes} represent ancestors to a certain subset of extant members, which we further define \emph{Sampling Nodes} and \emph{Transition Nodes}, whose differences will be clear in a moment.

For the sake of distinction and visualization, we define a set of \emph{colored ball}s, $(f,n)\in\Colors\times\N$, where $n$ is the label and $f$ is the color of the ball, and $\Colors\coloneqq\{\green,\black,\blue,\red,\purple\}$.
Therefore, we can define a genealogy \emph{node} as a quadruple $(n,t,w,z)$, where $n\in\N$ is the node's \emph{name}, $t\in\Time$ is its time, $w$ is the node's \emph{pocket} (\ie unordered pair of colored balls), and $z\in\N$ is the node's \emph{deme}. 
For a node $\Node{p}$, we will use $n(\Node{p})$, $t(\Node{p})$, $w(\Node{p})$, and $z(\Node{p})$ to denote the name, time, pocket, and deme of $\Node{p}$, respectively. 
We then assign black, green, blue, and purple balls to serve as pointers to the extant population at time $t$, to the internal nodes, to the samples, and to the transitions between demes, respectively.

@@

\paragraph{Effect of births, death, sampling and transistions}
Births, deaths, samples, and transitions change the topology of a structured genealogy. 
A death at time $t$ indicates the removal of a leaf in the genealogy and the dismiss of a node. The detailed definition can be found in \citet{King2021}.

A birth at time $t$ in a structured genealogy is similar to that in a unstructured one \citep{King2021}: a black ball is selected as the parent and a new node introduced as the child, who is in the same deme as the parent. Let $\Ball{b}$ be randomly selected black ball, named \emph{n}, which is the parent, and there exists a node $\Node{p}\in\Gen$ holding $\Ball{b}\in w(\Node{p})=\{\Ball{b},\Ball{b}'\}$ where $\Ball{b}'$ is the other ball held by $\Node{p}$. 
Here, we assume the parent is from Deme $i$ (\ie $n\in\Inv_i$) and the child is born into Deme $j$. 
We then can produce a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},i)$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, $\Ball{b}''=(\black,c)$, and update $\Inv_j=\Inv_j\cup\{c\}$. 
By swapping balls between $\Node{p}$ and $\Node{p}'$, it ends up that $w(\Node{p})=\{\Ball{g},\Ball{b}'\}$ and $w(\Node{p}')=\{\Ball{b},\Ball{b}''\}$. 
We denote the resulting sequence of nodes as $\mathrm{add}(\seq{\Gen},n)$.

A death at time $t$ is defined in the same way as that in \citet{King2021} and we continue to use the notation $\mathrm{drop}(\seq{\Gen},n)$ to denote the resulting sequence of nodes for a death, where an individual with name \emph{n} is selected randomly to be removed from the population.

A sample at time $t$ results in the participation of a new node with a new blue ball. 
The procedure is similar to the previous sampling in \citet{King2021} and the birth mentioned above: a black ball, named \emph{n}, where $n\in\Inv_j$, and held by $\Node{p}$ (\ie $w(\Node{p})=\{\Ball{b},\Ball{b}'\}$), is selected at random. 
Then a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},j)$ is generated, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\blue,q)$. 
At the end, $\Node{p}$ exchanges $\Ball{b}$ for $\Ball{g}$ with $\Node{p}'$. 
Here, $q$ is the name of the new blue ball, which is the ordinal number of the transition, $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is blue}\}|$. 
We use the notation $\mathrm{sample}(\seq{\Gen},n)$ as the resulting sequence of nodes.

A transition at time $t$ leads to the inclusion of a new node with a new purple ball. 
Let $\Ball{b}$ denote the
black ball with name \emph{n} and selected to indicate an extant in the population transits from his/her current
deme to a new one. 
Assuming this selected individual is from Deme $i$ (\ie $i\in\Inv_i$) and transiting to Deme $j$ ($j\neq i$), we then initialize a new node by taking $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},i)$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\purple,q)$. 
Note that, we update $\Inv_i=\Inv_i\setminus\{n\}$ and $\Inv_j=\Inv_j\cup\{n\}$. 
Similar to other events, we therefore swap $\Ball{b}$ for $\Ball{g}$ between nodes $\Node{p}$ and $\Node{p}'$ and insert $\Node{p}'$ at the last position of the node-sequence. 
Here, \emph{q} is the name of the new purple ball, which is the ordinal number of the transition,  $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is purple}\}|$. 
We call this new node $\Node{p}'$ the \emph{Transition Nodes} in the genealogy and denote the resulting sequence of nodes by $\mathrm{transition}(\seq{\Gen},n)$.

\paragraph{Event times}

Given a genealogy $\Gen$, the set of \emph{genealogical event times}, $\Eset(\Gen)$, is the set of all node times.
Its subsets, including those for different event types and those for different demes, are of interest.
In particular, we define
\begin{equation*}
  \begin{aligned}
    \Eset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})\ \text{is}\ j\}, \\
    \Masterset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, w(\Node{p})\ \text{contains a green ball}, z(\Node{p})\ \text{is}\ j\}, \\
    \Cset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, w(\Node{p})\ \text{contains two green balls}, z(\Node{p})\ \text{is}\ j\}, \\
    \Lset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, w(\Node{p})\ \text{contains no green balls}, z(\Node{p})\ \text{is}\ j\}, \\
    \Tset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, w(\Node{p})\ \text{contains a blue ball}, z(\Node{p})\ \text{is}\ j\}, \\
    \Zset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, w(\Node{p})\ \text{contains a purple ball}, z(\Node{p})\ \text{is}\ j\}, \\
    \Qset^j(\Gen) &\coloneqq \Masterset^j(\Gen) \cap \Tset^j(\Gen).
  \end{aligned}
\end{equation*}
With these definitions,
$\Masterset^j(\Gen)$ comprises the internal node times of nodes that are derived from nodes in Deme $j$,
$\Cset^j(\Gen)$ is the set of branch times of nodes that are produced by nodes in Deme $j$,
$\Lset^j(\Gen)$ contains the leaf times where the tips are in Deme $j$,
$\Tset^j(\Gen)$ holds the sample times for samples in Deme $j$,
$\Zset^j(\Gen)$ holds the transition times derived from nodes in Deme $j$,
and $\Qset^j(\Gen)$ is the set of \emph{direct-descent times} derived from samples in Deme $j$. Trivially, $\Masterset(\Gen) = \cup_j \Masterset^j(\Gen)$, $\Cset(\Gen) = \cup_j \Cset^j(\Gen)$, $\Lset(\Gen) = \cup_j\Lset^j(\Gen)$, $\Tset(\Gen)=\cup_j\Tset^j(\Gen)$, $\Zset(\Gen)=\cup_j\Zset^j(\Gen)$, and $\Qset(\Gen)=\cup_j\Qset^j(\Gen)$, representing all internal node times, branch times, leaf times, sample times, transistion times, and direct-descent times for $\Gen$, respectively.

@@

\fontsize{10pt}{12pt}\selectfont
\begin{equation}\label{eq:Gdef}
  \mathrm{geneal}(\omega)\coloneqq
  \begin{cases}
    \left(k,0,\{(\green,k),(\black,k)\},0\right)_{k=0}^{I(\X_0(\omega))}, &\text{if}\ K(\omega)=0;\\
    \mathrm{add}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Bset;\\
    \mathrm{drop}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Dset;\\
    \mathrm{sample}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Sset;\\
    \mathrm{transition}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Mset;\\
    \mathrm{geneal}(\predec{\omega}), &\text{otherwise};
  \end{cases}
\end{equation}
\normalsize

@@

\paragraph{Branch color}

>>>>

\begin{figure}
  \resizebox{0.6\linewidth}{!}{
    \begin{tikzpicture}[scale=1.8]
      \usetikzlibrary{arrows}
      \tikzstyle{coordinate}=[inner sep=0pt,outer sep=0pt]
      \tikzstyle{det}=[color=black, thick, >=stealth]
      \tikzstyle{stoch}=[color=black, dashed, thick, >=stealth]
      \node (A) at (0,0) {$\Master_t$};
      \node (H) at (30:1) {$\Hist_t$};
      \node (X) at (30:2) {$\X_t$};
      \node (G) at (-30:1) {$\Gen_t$};
      \node (I) at (-30:2) {$\Inv_t$};
      \node (P) at (0:1.732) {$\Prune_t$};
      \node (O) at (1.732,.5) {$\Och_i$};
      \node (V) at (0:2.732) {$\Vis_t$};
      \node (W) at (2.732,.5) {$\Wch_i$};
      \draw [det,->] (A) -- (H);
      \draw [det,->] (H) -- (X);
      \draw [det,->] (A) -- (G);
      \draw [det,->] (G) -- (I) node[midway,below,sloped] {$\mathrm{ext}$};
      \draw [det,->] (G) -- (P) node[midway,below,sloped] {$\mathrm{prune}$};
      \draw [det,->] (P) -- (V) node[midway,below,sloped] {$\mathrm{obs}$};
      \draw [det,->] (P) -- (O);
      \draw [det,->] (V) -- (W);
      \draw [stoch,->] (H) -- (G);
    \end{tikzpicture}
  }
  \caption{
    \label{fig:constellation}
    Relations among the various Markov processes discussed in the paper.
    Deterministic maps are indicated with solid arrows;
    random maps are shown as dashed arrows.
    All the maps shown commute.
    $\X_t$ is the \emph{population process}, a model of the dynamics of some system, which we take as a starting point.
    $\Hist_t$ is the \emph{history process}, which records the full history of $\X_t$.
    $\Inv_t$ is the \emph{inventory process}: at each time $t$, $\Inv_t$ is an inventory of all extant individuals in the population, each of which has a globally unique name.
    $\Gen_t$ is the \emph{genealogy process}, which captures the precise genealogical relationships among all individuals in $\Inv_t$, as well as among any samples that have been taken from the population.
    $\Vis_t$ is the \emph{visible genealogy process}, which is $\Gen_t$ pruned so that only relationships among samples remain.
    Finally $\Wch_i$ is the \emph{embedded chain of the visible genealogy process}, which is $\Vis_{s_i}$, $s_i$ being the time of the $i$-th sample.
    All of these processes can be obtained via deterministic procedures applied to the \emph{master process} $\Master_t$, as described in the text.
  }
\end{figure}

@@

\paragraph{Migrations}
Infective demes are interchangeable and we define the migration function from Deme $i$ to $j$ for any $i\neq j$ and $i,j=1,\dots,J$ as $M_{ij}:\Uspace\to\N$ to represent the number of infectives migrating to Deme $j$ from Deme $i$.
Thus the set of migration is $\Mset_{ij}\coloneqq\,M^{-1}_{ij}(\{1\})$ since we assume single migration at a single migration event. 
Similarly, we disallow coninciding migration events and other events.

@@

\paragraph{Deaths and population size}
The population, the set of infected individuals in the context of epidemiology, is structured, categorized into interchangeable demes. 
We can then define population size function and death functions in Deme $j$, respectively, as $I^{j}, D_{j}:\Xspace\to\N$ for any $j=1,\cdots,J$. 
\begin{equation}
  \jumprate_u(t,x) > 0 \Longrightarrow I^j(x+u)-I^j(x)=\sum_{i}\,B_{ij}(u)+\sum_{i}M_{ij}(u)-D_j(u)-\sum_{i}M_{ji}(u),
\end{equation}

for all $x, u\in\Xspace$. We disallow simultaneous death at a single death event at this stage and define the set of death events in Deme $j$ as $\Dset_j\coloneqq\,D^{-1}(\{1\})$. 
We also insist, birth and death events don't coincide for any demes, $i,j$, that is,  $\Bset_{j}\cap\Dset_{i}=\emptyset$. 
Eventually, we can have the function for total population size by letting $I(x)\coloneqq\sum_{j=1}^J\,I^j(x)$ and it's trivial that

@@

We can customize the state and event vector to fit a complex system. 
In SI$^2$R model, we have two different while interchangeable demes of infectious individuals, the \emph{per capita} transmissibility therefore in the first deme is $b_1(t)$ whilst in the second deme is $b_2(t)$. 
Migration rates between Deme $i$ and $j$ are also defined: Deme $i$ infection turns into Deme $j$ at rate $r_{ij}$ for $i,j=1,2$ and $i\neq j$. Furthermore, we also suppose that, once being infected, a susceptible individual becomes Deme 1 infectious with probability $\rho$ while Deme 2 infectious with probability $1-\rho$. 
To specify each event and the number of individuals involved, we define a relatively complex state vector by setting $d=10$ and $\sg{X}=\left(s,i_1,i_2,b_{11},b_{12},b_{21},b_{22},m_{12},m_{21},g\right)$, for $s$, $i_1$, and $i_2$ being the population size of the susceptible, the infectives in Deme 1, and those in Deme 2, respectively, and $b_{11}$, $b_{12}$, $b_{21}$, $b_{22}$, $m_{12}$, and $m_{21}$ being the cumulative number of events of parents in Deme 1 giving to children in Deme 1 and 2, parents in Deme 2 giving birth to children in Deme 1 and 2, individuals migrating from Deme 1 to Deme 2, and vice versa, and sampling, respectively. 
In this case, we can summarize the event rates as:

@@

The nine non-zero rates above represent those of infection into Deme 1 infected by individuals in Deme 1, infection into Deme 2 infected by individuals in Deme 1, infection in Deme 1 infected by individuals in Deme 2, infection into Deme 2 infected by individuals in Deme 1, migration from Deme 1 to Deme 2, migration from Deme 2 to Deme 1, recovery from Deme 1, recovery from Deme 2, and sampling, respectively. 
We then summarize the population functions and sets as follows:

@@

\paragraph{Migration}
One special case of the SI$^2$R model is that with one-time migration or introduction. 
In other words, an infectious individual from another population join our interested population at a certain point of time $\tau$ and then stay and spread the disease. 
In this case, if this specific individual is defined as Deme 1 while other infectious individuals as Deme 2, the migration rate $r_{12}$ is a Dirac delta function $\delta(t)$, where $\delta(t)=\infty$ for $t=\tau$ and $\delta(t)=0$ for $t\neq\tau$, and $r_{21}=0$ all the time.

@@

\begin{equation}
  \mathrm{inven}(\omega,j)\coloneqq
  \begin{cases}
    \mathrm{initialize}(I^j(\sg{X}_0(\omega)), & \text{if }K(\omega)=0; \\
    \mathrm{add}(\mathrm{inven}(\predec{\omega},j)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\Bset_j;\\
    \mathrm{drop}(\mathrm{inven}(\predec{\omega},j),N_{K(\omega)}(\omega)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\Dset_j; \\
    \mathrm{movein}(\mathrm{inven}(\predec{\omega},j),N_{K(\omega)}(\omega)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\bigcup_i\Mset_{ij}; \\
    \mathrm{drop}(\mathrm{inven}(\predec{\omega},j),N_{K(\omega)}(\omega)), & \text{if }K(\omega)>0\text{ and }U_K(\omega)\in\bigcup_i\Mset_{ji}; \\
    \mathrm{inven}(\predec{\omega},j), & \text{otherwise}.
  \end{cases}
\end{equation}

@@

\paragraph{Structured genealogies}
Types of nodes are defined in a structured genealogy: \emph{Leaves} are the extant members at current time and \emph{Internal Nodes} represent ancestors to a certain subset of extant members, which we further define \emph{Sampling Nodes} and \emph{Migration Nodes}, whose differences will be clear in a moment.

For the sake of distinction and visualization, we define a set of \emph{colored ball}s, $(f,n)\in\Colors\times\N$, where $n$ is the label and $f$ is the color of the ball, and $\Colors\coloneqq\{\green,\black,\blue,\red,\purple\}$.
Therefore, we can define a genealogy \emph{node} as a quadruple $(n,t,w,z)$, where $n\in\N$ is the node's \emph{name}, $t\in\Time$ is its time, $w$ is the node's \emph{pocket} (\ie unordered pair of colored balls), and $z\in\N$ is the node's \emph{deme}. 
For a node $\Node{p}$, we will use $n(\Node{p})$, $t(\Node{p})$, $w(\Node{p})$, and $z(\Node{p})$ to denote the name, time, pocket, and deme of $\Node{p}$, respectively. 
We then assign black, green, blue, and purple balls to serve as pointers to the extant population at time $t$, to the internal nodes, to the samples, and to the migrations between demes, respectively.

@@

\paragraph{Effect of births, death, sampling and migration}
Births, deaths, samples, and migrations change the topology of a structured genealogy. 
A death at time $t$ indicates the removal of a leaf in the genealogy and the dismiss of a node. The detailed definition can be found in \citet{King2021}.

A birth at time $t$ in a structured genealogy is similar to that in a unstructured one \citep{King2021}: a black ball is selected as the parent and a new node introduced as the child, who is in the same deme as the parent. Let $\Ball{b}$ be randomly selected black ball, named \emph{n}, which is the parent, and there exists a node $\Node{p}\in\Gen$ holding $\Ball{b}\in w(\Node{p})=\{\Ball{b},\Ball{b}'\}$ where $\Ball{b}'$ is the other ball held by $\Node{p}$. 
Here, we assume the parent is from Deme $i$ (\ie $n\in\Inv_i$) and the child is born into Deme $j$. 
We then can produce a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},i)$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, $\Ball{b}''=(\black,c)$.
By swapping balls between $\Node{p}$ and $\Node{p}'$, it ends up that $w(\Node{p})=\{\Ball{g},\Ball{b}'\}$ and $w(\Node{p}')=\{\Ball{b},\Ball{b}''\}$. 
We denote the resulting sequence of nodes as $\mathrm{add}(\seq{\Gen},n)$.

% A death at time $t$ is defined in the same way as that in \citet{King2021} and we continue to use the notation $\mathrm{drop}(\seq{\Gen},n)$ to denote the resulting sequence of nodes for a death, where an individual with name \emph{n} is selected randomly to be removed from the population.

A death at time $t$ is quite different from that defined in \citet{King2021}, while we continue to use the notation $\mathrm{drop}(\seq{\Gen},n)$ to denote the resulting sequence of nodes for a death, where an individual with name \emph{n} is selected randomly to be removed from the current population. 
When a death occurs, a black ball $\Ball{b}$ with name \emph{n} is selected arbitrarily, whose holder is node $\Node{p}$, \ie $\Ball{b}\in w(\Node{p})=\{\Ball{b},\Ball{b}'\}$. 
There exists an unique $\green$ ball and an unique node $\Node{p}'$ such that $\Ball{g}=(\green, n(p))$ and $\Ball{g}\in w(\Node{p}')=\{\Ball{g},\Ball{b}''\}$. 
The $\mathrm{drop}(\seq{\Gen},n)$ operation then proceeds differently, conditioned on the color and the deme of the other ball, $\Ball{b}'$, held by $\Node{p}$: 
\begin{inparaenum}[(1)]
  \item if $\Ball{b}'$ is $\green$, swap $\Ball{g}$ and $\Ball{b}'$ between $\Node{p}$ and $\Node{p}'$ then remove $\Node{p}$ from the sequence of nodes;
  \item if $\Ball{b}'$ is $\black$, the above procedure is conducted when $\Ball{b}'$ and $\Node{p}$ are in the same deme, otherwise, replace $\Ball{b}$ by a new purple ball $\Ball{c}=\{\purple, q\}$ where $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is purple}\}|$;
  \item if $\Ball{b}'$ is $\blue$, replace $\Ball{b}$ by a new red ball with the name of $\Ball{b}'$;
  \item if $\Ball{b}'$ is $\purple$, change the deme of $\Ball{b}$ to that of $\Node{p}$, swap $\Ball{g}$ and $\Ball{b}$, and delete $\Node{p}$.
\end{inparaenum}
By doing aforementioned steps, we can keep track of the sampled individuals and the migration at a birth.

A sample at time $t$ results in the participation of a new node with a new blue ball. 
The procedure is similar to the previous sampling in \citet{King2021} and the birth mentioned above: a black ball, named \emph{n}, where $n\in\Inv_j$, and held by $\Node{p}$ (\ie $w(\Node{p})=\{\Ball{b},\Ball{b}'\}$), is selected at random. 
Then a new node $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},j)$ is generated, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\blue,q)$. 
At the end, $\Node{p}$ exchanges $\Ball{b}$ for $\Ball{g}$ with $\Node{p}'$. 
Here, $q$ is the name of the new blue ball, which is the ordinal number of the migration, $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is blue}\}|$. 
We use the notation $\mathrm{sample}(\seq{\Gen},n)$ as the resulting sequence of nodes.

A migration at time $t$ leads to the inclusion of a new node with a new purple ball. 
Let $\Ball{b}$ denote the
black ball with name \emph{n} and selected to indicate an extant in the population migrates from his/her current
deme to a new one. 
Assuming this selected individual is from Deme $i$ (\ie $i\in\Inv_i$) and migrating to Deme $j$ ($j\neq i$), we then initialize a new node by taking $\Node{p}'=(c,t,\{\Ball{g},\Ball{b}''\},i)$, where $c=\mathrm{next}(\Names)$, $\Ball{g}=(\green,c)$, and $\Ball{b}''=(\purple,q)$. 
Note that, we update $\Inv_i=\Inv_i\setminus\{n\}$ and $\Inv_j=\Inv_j\cup\{n\}$. 
Similar to other events, we therefore swap $\Ball{b}$ for $\Ball{g}$ between nodes $\Node{p}$ and $\Node{p}'$ and insert $\Node{p}'$ at the last position of the node-sequence. 
Here, \emph{q} is the name of the new purple ball, which is the ordinal number of the migration,  $q=|\{\Ball{b}\in w(\Gen)\;|\;\Ball{b}\ \text{is purple}\}|$. 
We call this new node $\Node{p}'$ the \emph{Migration Nodes} in the genealogy and denote the resulting sequence of nodes by $\mathrm{migrate}(\seq{\Gen},n)$.

\paragraph{Genealogical event times}

\QYL{the definition for sets of event times is problematic.}

Given a genealogy $\Gen$, the set of \emph{genealogical event times}, $\Eset(\Gen)$, is the set of all node times.
Its subsets, including those for different event types and those for different demes, are of interest.
In particular, we define
\begin{equation*}
  \begin{aligned}
    \Eset^j(\Gen) &\coloneqq \{t(\Node{p})\;|\;\Node{p}\in\Gen, z(\Node{p})\ \text{is}\ j\}, \\
    \Masterset^j(\Gen) &\coloneqq , \\
    \Cset^{j,j'}(\Gen) &\coloneqq , \\
    \Lset^j(\Gen) &\coloneqq , \\
    \Tset^j(\Gen) &\coloneqq , \\
    \Zset^{j,j'}(\Gen) &\coloneqq , \\
    \Qset^j(\Gen) &\coloneqq \Masterset^j(\Gen) \cap \Tset^j(\Gen).
  \end{aligned}
\end{equation*}
With these definitions,
$\Masterset^j(\Gen)$ comprises the internal node times of nodes that are derived from nodes in Deme $j$,
$\Cset^{j,j'}(\Gen)$ is the set of coalescent times of nodes in Deme $j$ and containing a ball in Deme $j'$,
$\Lset^j(\Gen)$ contains the leaf times where the tips are in Deme $j$,
$\Tset^j(\Gen)$ holds the sample times for samples in Deme $j$,
$\Zset^{j,j'}(\Gen)$ holds the deme-transition times derived from nodes in Deme $j$ and containing a ball in Deme $j'$,
and $\Qset^j(\Gen)$ is the set of \emph{direct-descent times} derived from samples in Deme $j$.

@@

\fontsize{10pt}{12pt}\selectfont
\begin{equation}\label{eq:Gdef}
  \mathrm{geneal}(\omega)\coloneqq
  \begin{cases}
    \left(k,0,\{(\green,k),(\black,k)\},0\right)_{k=0}^{I(\X_0(\omega))}, &\text{if}\ K(\omega)=0;\\
    \mathrm{add}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Bset;\\
    \mathrm{drop}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Dset;\\
    \mathrm{sample}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Sset;\\
    \mathrm{migrate}(\mathrm{geneal}(\predec{\omega}),N_{K(\omega)}(\omega)), &\text{if}\ K(\omega)>0\ \text{and}\ U_{K(\omega)}\in\Mset;\\
    \mathrm{geneal}(\predec{\omega}), &\text{otherwise};
  \end{cases}
\end{equation}
\normalsize

@@

\paragraph{Visible genealogy process}
The sampled genealogy contains only relations among genomic samples. 
In other words, information in a full genealogy about unsampled individuals and deme-transitions are unobserved in a sampled genealogy, which we call it a visible genealogy. 
We divide the procedure of deriving a visible genealogy from a full genealogy into two separate steps: \emph{pruning} and \emph{obscuring}. 
The former drops all unsampled individuals (\ie black balls) in a genealogy and remains the relationship between samples and deme-transition records of samples, and the latter further erases all deme-transition records of samples and obtains an obscured genealogy.
